---
title: "LabAssignment1"
author: "Tian Fu"
date: "11/1/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(expm)
```

# Problem 1
```{r}
# load data
markov = read.table("~/Desktop/MSiA400/markov100.txt", head=FALSE)
P = as.matrix(markov) # transition probability matrix
```
```{r}
names = c(1:100) # assign names to each state, from 1 to 100
colnames(P) = names
rownames(P) = names
```
# 1(a)
```{r}
a <- rep(0, 100) # initial distribution
a[1] <- 1 # at State 1 with probability one
```
Since we are at State 1 now, in the initial distribution a, we are at State 
1 with probability one.

```{r}
prob5 <- a %*% (P%^%10)
prob5
prob5[1,5]
```
Given we are at State 1 now, the probability of being in State 5 after
10 transitions is:

$p_{1,5}(10)$ = 0.045091

# 1(b)
```{r}
a2 <- rep(0, 100)
a2[1:3] <- 1/3
```
Since we are at one of States 1,2, and 3 with equal probabilities, the
initial probability of being in one of these three is $\frac{1}{3}$

```{r}
prob10 <- a2 %*% (P%^%10)
prob10
prob10[1,10]
```
Given we are initially at State 1,2,3 with equal probability,
the probability of being in State 10 after 10 transitions is:

$p_{i,10}(10)$ = 0.08268901

where i=1,2,3 with equal probability

# 1(c)
```{r}
Q = t(P)-diag(100)
Q[100,] = rep(1,100) # replace the last row with all ones
rhs = rep(0,100)
rhs[100] = 1
Pi = solve(Q) %*% rhs
Pi # steady state probabilities
```

As shown above, the steady state probability of being in State 1 is:
$\pi_1$ = 0.0125658938

# 1(d)
```{r}
B = P[1:99,1:99] # exclude destination State 100
Q2 = diag(99) - B
e = rep(1,99)
m = solve(Q2) %*% e
m # mean first passage time from State i to State 100, i != 100
```

From above, the mean first passage time from State 1 to State 100 is:
$m_{1,100}$ = 254.939463

# Problem 2
```{r}
# load data
webtraffic <- read.table("~/Desktop/MSiA400/webtraffic.txt", head=TRUE)
```

# 2(a)
```{r}
columnSum <- unname(colSums(webtraffic))
Traffic <- t(matrix(columnSum, nrow = 9, ncol = 9))
Traffic
```

From above, matrix Traffic that counts total traffic between State i to State j for all i = 1,2,3,4,5,6,7,8,9 and j = 1,2,3,4,5,6,7,8,9 is:

\[
 Traffic =
 \begin{bmatrix}{}
  0 & 447 & 553 & 0 & 0 & 0 & 0 & 0 & 0 \\ 
  0 & 23 & 230 & 321 & 0 & 0 & 0 & 0 & 63 \\
  0 & 167 & 43 & 520 & 0 & 0 & 0 & 0 & 96 \\
  0 & 0 & 0 & 44 & 158 & 312 & 247 & 0 & 124 \\
  0 & 0 & 0 & 0 & 22 & 52 & 90 & 127 & 218 \\ 
  0 & 0 & 0 & 0 & 67 & 21 & 0 & 294 & 97 \\
  0 & 0 & 0 & 0 & 0 & 94 & 7 & 185 & 58 \\
  0 & 0 & 0 & 0 & 262 & 0 & 0 & 30 & 344 \\
  0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 \end{bmatrix}
\]

# 2(b)
```{r}
Traffic[9,1] <- 1000

for (i in 1:nrow(Traffic)){
  rowSum = sum(Traffic[i,])
  Traffic[i,] = Traffic[i,] / rowSum
}

Traffic # P
```

One step transition probability matrix is:

\[
 P =
 \begin{bmatrix}{}
  0 & 0.44700000 & 0.55300000 & 0 & 0 & 0 & 0 & 0 & 0 \\ 
  0 & 0.03610675 & 0.36106750 & 0.50392465 & 0 & 0 & 0 & 0 & 0.0989011 \\
  0 & 0.20217918 & 0.05205811 & 0.62953995 & 0 & 0 & 0 & 0 & 0.1162228 \\
  0&0&0&0.04971751 & 0.1785311 & 0.35254237 & 0.27909605 & 0 & 0.1401130 \\
  0&0&0&0&0.0432220 & 0.10216110 & 0.17681729 & 0.24950884 & 0.4282908 \\ 
  0 & 0 & 0 & 0 & 0.1398747 & 0.04384134 & 0 & 0.61377871 & 0.2025052 \\
  0 & 0 & 0 & 0 & 0 & 0.27325581 & 0.02034884 & 0.53779070 & 0.1686047 \\
  0 & 0 & 0 & 0 & 0.4119497 & 0 & 0 & 0.04716981 & 0.5408805 \\
  1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 \end{bmatrix}
\]

# 2(c)
```{r}
Q3 = t(Traffic)-diag(9)
Q3[9,] = rep(1,9)
rhs2 = rep(0,9)
rhs2[9] = 1
Pi2 = solve(Q3) %*% rhs2
Pi2
```

From above, the steady state probability vector is:

\[
 \pi =
 \begin{bmatrix}{}
 0.15832806 \\
 0.10085497 \\
 0.13077897 \\
 0.14012033 \\
 0.08058898 \\
 0.07583914 \\
 0.05446485 \\
 0.10069664 \\
 0.15832806 \\
 \end{bmatrix}
\]

# 2(d)
```{r}
avg_time <- c(0.1,2,3,5,5,3,3,2,0)
expected <- t(Pi2) %*% avg_time
expected
```

The average time a visitor spend on a page is 2.305731 minutes

```{r}
B2 = Traffic[1:8,1:8] # exclude destination State 9
Q_m = diag(8) - B2
e2 = rep(1,8)
m2 = solve(Q_m) %*% e2
m2 # mean first passage time from State i to State 9, i != 9
```

From above, the mean first passage time from Page 1 to Page 9 is:
$m_{1,9}$ = 5.316

Therefore, average time a visitor spend on the website (until she leaves)
is:

$5.316$*$2.305731$ = $12.25727$ minutes

# 2(e)

In the output 2(c), Pages 3 and 4 have higher values than others, excluding
Pages 1 and 9.
```{r}
Traff <- t(matrix(columnSum, nrow = 9, ncol = 9))
Traff # Traffic matrix
```

```{r}
two_to_three <- Traff[2,3] #current outgoing traffic to State3 from State2
two_to_four <- Traff[2,4] #current outgoing traffic to State4 from State2
# new assignments after linking Page 2 to 6,7
Traff[2,3] <- two_to_three*0.7
Traff[2,6] <- two_to_three*0.3
Traff[2,4] <- two_to_four*0.8
Traff[2,7] <- two_to_four*0.2
Traff
```
### calculate new steady state probability
```{r}
Traff[9,1] <- 1000
# calculate one step transtion probability matrix and stored in Traff
for (i in 1:nrow(Traff)){
  rowSum = sum(Traff[i,])
  Traff[i,] = Traff[i,] / rowSum
}

Q4 = t(Traff)-diag(9)
Q4[9,] = rep(1,9)
rhs3 = rep(0,9)
rhs3[9] = 1
Pi3 = solve(Q4) %*% rhs3
Pi3
```

The new steady state probability vector is:

\[
 \pi_2 =
 \begin{bmatrix}{}
 0.16162840 \\
 0.10034341 \\
 0.12104331 \\
 0.12275720 \\
 0.08164613 \\
 0.08250884 \\
 0.06003218 \\
 0.10841213 \\
 0.16162840 \\
 \end{bmatrix}
\]

Comparing with original $\pi$, after creating new links from Page 2 to Page
6,7, steady state probabilities in Page 3,4 decrease and steady state
probability in Page 2 slightly decreases. Steady state probabilities in
other Pages all increase.

```{r}
# variance of pi before change
var(Pi2)
# variance of pi after change
var(Pi3)
```

From above, $var(\pi)$ = 0.001410675 > $var(\pi_2)$ = 0.001219604

As the variance of steady state probability decreases after creating new
links, there is less variation after introducing the new links.
So, the link helped balancing the traffic.